Prognostic and biological relevance of collagen-related genes in prostate cancer

### 2025-02-21 paper/markdown/manuscript_figures_tables.Rmd
Supplementary Figures
Supplementary figure legends

_Prostate cancer (PCA) patients in the pooled GEO (total: n = `r uni_globals$n_numbers$n_total["geo_pool"]`, biochemical relapse [BCR]: n = `r uni_globals$n_numbers$n_events["geo_pool"]`), TCGA (total: n = `r uni_globals$n_numbers$n_total["tcga"]`, BCR: n = `r uni_globals$n_numbers$n_events["tcga"]`), and DKFZ cohort (total: n = `r uni_globals$n_numbers$n_total["dkfz"]`, BCR: n = `r uni_globals$n_numbers$n_events["dkfz"]`) were classified as high and low expressors for each of the `r `length(globals$genes)` collagen-related transcripts with the expression cutoffs corresponding to the largest differences in survival between the high and low expressors._
_BCR risk was compared between the high and low expressors for each gene by univariable Cox proportional hazard regression. P values were corrected for multiple testing with the false discovery rate (FDR) method._
_Standalone BCR risk markers were defined as transcripts associated with unfavorable prognosis (high versus low expressors, hazard ratio [HR] > 1, pFDR < 0.05) or favorable prognosis (HR < 1, pFDR < 0.05) in all three cohorts._

@@

_(B) BCR-free survival in PCA patients stratified by tertiles of the predictor scores of the best performing GBM model. Statistical significance of differences between the predictor score tertiles was determined by Peto-Peto test corrected for multiple testing wit the false discovery rate method. Fractions of BCR-free patients and minimum/maximum scaled time after diagnosis are visualized in Kaplan-Meier plots. Numbers of patients and BCR in the predictor score tertiles are indicated in the legends. P values are displayed in the plots._

_(C) Importance of the explanatory variables for BCR-risk prediction by the GBM model was gauged by difference in sum of squared error ($\Delta SSE$) attributed to inclusion of a particular collagen-related gene in the GBM learner ensemble. High $\Delta SSE$ values are characteristic for highly influential variables. $\Delta SSE$ values for the variables that contributed substantially to the BCR risk prediction ($\Delta SSE$ > 0) are presented in a bar plot._

>>>>

_Prostate cancer (PCA) patients in the pooled GEO (total: n = `r uni_globals$n_numbers$n_total["geo_pool"]`, biochemical relapse [BCR]: n = `r uni_globals$n_numbers$n_events["geo_pool"]`), TCGA (total: n = `r uni_globals$n_numbers$n_total["tcga"]`, BCR: n = `r uni_globals$n_numbers$n_events["tcga"]`), and DKFZ cohort (total: n = `r uni_globals$n_numbers$n_total["dkfz"]`, BCR: n = `r uni_globals$n_numbers$n_events["dkfz"]`) were classified as high and low expressors for each of the `r `length(globals$genes)` collagen-related transcripts with the expression cutoffs corresponding to the largest differences in survival between the high and low expressors._
_BCR risk was compared between the high and low expressors for each transcript by univariable Cox proportional hazard regression. P values were corrected for multiple testing with the false discovery rate (FDR) method._
_Standalone BCR risk markers were defined as transcripts associated with unfavorable prognosis (high versus low expressors, hazard ratio [HR] > 1, pFDR < 0.05) or favorable prognosis (HR < 1, pFDR < 0.05) in all three cohorts._

@@

_(B) BCR-free survival of PCA patients stratified by tertiles of the predictor scores of the best performing GBM model. Statistical significance of differences between the predictor score tertiles was determined by Peto-Peto test corrected for multiple testing wit the false discovery rate method. Fractions of BCR-free patients and minimum/maximum scaled time after diagnosis are visualized in Kaplan-Meier plots. Numbers of patients and BCR in the predictor score tertiles are indicated in the legends. P values are displayed in the plots._

_(C) Importance of the explanatory variables for BCR-risk prediction by the GBM model was gauged by difference in sum of squared error ($\Delta SSE$) attributed to inclusion of a particular collagen-related transcript in the GBM learner ensemble. High $\Delta SSE$ values are characteristic for highly influential variables. $\Delta SSE$ values for the variables that contributed substantially to the BCR risk prediction ($\Delta SSE$ > 0) are presented in a bar plot._
